\section{Horizontal Agreements}

\subsection{Overview}

\subsubsection{Cartel Basics}

\begin{enumerate}
    \item Three kinds of market allocation:
    \begin{enumerate}
        \item Territory.
        \item Customer.
        \item Time (e.g., bid rigging).
    \end{enumerate}
    \item Market allocation is more harmful than price fixing. It rules out 
    \emph{all} competition and is easier to coordinate.
    \item Our economy encourages vertical agreements but views horizontal 
    agreements with suspicion.
    \item Sherman Act \S\ 1 outlaws every unreasonable agreement in restraint of 
    trade.
    \item ``Unreasonable'' means that the anticompetitive effects outweigh the 
    procompetitive efficiencies.
    \item So: today, \textbf{\S\ 1 outlaws every agreement for which AE > PE}.
\end{enumerate}

\subsubsection{Horizontal Agreements}

\begin{enumerate}
    \item \textbf{Per se illegal}: no need to show AE > PE. Price fixing, market 
    division, group boycotts.
    \item Other cases apply the \textbf{rule of reason} analysis, which asks 
    only if AE > PE.
\end{enumerate}

\subsection{Hardcore Cartels and Efficient Collaborations}

\subsubsection{\emph{United States v. Andreas}}

\begin{enumerate}
    \item Lysine price fixing: a classic example of a per se illegal agreement.
\end{enumerate}

\subsubsection{False Positives and the Ancillary Restraints Analysis: 
\emph{Broadcast Music, Inc. v. Columbia Broadcasting System, Inc.}}

\begin{enumerate}
    \item ASCAP literally required price fixing by a cartel of composers.
    \item But that fixed price was absolutely necessary for the venture to work.  
    Without it, transaction costs would be excessive.
    \item So even though this could have been a per se offense, the court 
    applied the rule of reason and held for BMI.
    \item \textbf{Courts will only apply the per se treatment if there has been 
    significant judicial experience in the area.}
    \item \textbf{Ancillary restraints analysis}: the restrictive agreement here 
    was \textbf{ancillary} to an underlying transaction, so the court analyzed 
    it under the rule of reason.
\end{enumerate}

\subsection{Analyzing a Horizontal Agreement Problem}

\begin{enumerate}
    \item Was this a \textbf{per se offense} (price fixing, market allocation, 
    boycott)?
    \begin{enumerate}
        \item Yes:
        \begin{enumerate}
            \item Has there been significant judicial experience?
            \begin{enumerate}
                \item Yes: \emph{violation of \S\ 1}.
                \item No: apply rule of reason.
            \end{enumerate}
        \end{enumerate}
        \item No: go to rule of reason analysis.
    \end{enumerate}
    \item Was the agreement \textbf{ancillary} to an underlying procompetitive 
    transaction (as in BMI)?
    \begin{enumerate}
        \item Yes: apply rule of reason.
        \item No: \emph{violation of \S\ 1}.
    \end{enumerate}
    \item \textbf{Rule of reason}:
    \begin{enumerate}
        \item Is there a plausible theory of harm? I.e., is the defendant's 
        market share greater than 30\%?
        \item Is there proof of harm?
        \item Court asks: AE > PE?
    \end{enumerate}
\end{enumerate}

\subsection{New Business Models}

\subsubsection{Upstream Markets: \emph{United States v. Adobe Systems, Inc.}}

\begin{enumerate}
    \item The DOJ alleged that six companies---Adobe, Apple, Google, Intel, 
    Intuit, and Pixar---entered into no cold-call agreements to solicit 
    employees.
    \item The DOJ alleged the agreements were \textbf{per se unlawful under \S\ 
    1}.
    \item (The DOJ would likely fail on a rule of reason case because the 
    plaintiffs lack market share and the agreements had plausible procompetitive 
    justifications. The DOJ was trying to get the companies on a technicality.)
\end{enumerate}

\subsubsection{Ebooks: \emph{United States v. Apple, Inc.}}

Apple's ebook deals with publishers were per se violations of \S\ 1.

\begin{enumerate}
    \item The DOJ and 33 states claimed Apple and five book publishing companies 
    conspired to fix ebook prices in violation of \S\ 1 of the Sherman Act.
    \item The publishers settled.
    \item Facts:
    \begin{enumerate}
        \item The publishers were unhappy about Amazon's \$9.99 price for 
        ebooks.\footnote{p. 11.}
        \item Before April 2010, all publishers used the \textbf{wholesale  
        model}, in which the publisher sets a list price (i.e., a suggested 
        retail price) and then sells the books to a retailer at a wholesale 
        price (usually a percentage of the list price). The retailer then sells 
        the book at whatever price it chooses.\footnote{p. 14.} Amazon's \$9.99 
        was equal to or less than wholesale price of many of its ebooks, meaning 
        it was often selling ebooks as loss leaders.\footnote{pp. 15--17.}
        \item In 2009, the publishers started \textbf{windowing} (delaying ebook 
        publication after the release of the printed version). The publishers 
        synchronized the announcement of their windowing strategies.\footnote{p.  
        21.} But they believed windowing was not a long-term solution to 
        Amazon's \$9.99 price because it promoted piracy and penalized ebook 
        customers.
        \item Apple planned to announce the iPad on January 27, 2010. It wanted 
        to include an iBookstore. It assumed it would buy ebooks under the 
        wholesale model. Eddy Cue, SVP at Apple, began negotiating with the Big 
        Six.\footnote{p. 32.}
        \item Cue proposed an \textbf{agency model} in which the publishers 
        would set the price and Apple would earn a percentage of each sale.  
        Apple also required \textbf{price caps}.\footnote{p. 39.}
        \item Apple realized that it would be at a disadvantage if Amazon's 
        agreements with the publishers continued to use the wholesale model, 
        because Amazon could continue to undercut Apple's prices.  ``To ensure 
        that the iBookstore would be competitive at higher prices, Apple 
        concluded that it needed to eliminate all retail price competition.  
        Thus, the final component of its agency model required the Publishers to 
        move all of their e-tailers to agency.''\footnote{pp. 39--40.}
        \item Specifically, Cue proposed an agency model with a 30\% margin for 
        Apple, price caps, and a requirement that the Publishers switch all of 
        their other ebook agreements to the agency model.\footnote{p. 42.}
        \item Apple realized it could use a \textbf{most-favored nation (MFN) 
        clause} to force the publishers to adopt agency agreements with Amazon.  
        The MFN clause guaranteed that Apple's ebook retail prices would be the 
        lowest on the market. Amazon could continue to sell ebooks at \$9.99, 
        but Apple could match the price (and still earn 30\%). So, if the 
        publishers wanted to accomplish their goal of raising ebook prices 
        market-wide, they would have to adopt the agency model with Amazon to 
        force Amazon to raise its prices.\footnote{pp. 47--48.}
        \item The agency model would always be less profitable for publishers 
        than the wholesale model. (For instance, say Amazon and Apple both sell 
        the ebook at retail for \$12.99. Under the wholesale agreement with 
        Amazon in which the publisher's wholesale price to Amazon is \$13, the 
        publisher earns \$13 per book. But under the agency agreement with 
        Apple, the publisher would only receive 70\% of the revenue, or \$9.10.) 
        So the publishers would not have adopted the agency model unless it 
        suited their long-term goal of raising prices market-wide.\footnote{pp. 
        53--54.}
        \item Apple secured contracts with four of the Big Six publishers, all 
        of which soon told Amazon that they wanted to switch to the agency 
        model.\footnote{p. 69, 73.} HarperCollins was the last holdout. Steve 
        Jobs convinced James Murdoch (of News Corp, which owned HC) that
        Apple would help the publishers form a united front against Amazon (``so 
        there is no leap of faith here''\footnote{p. 78.}). HC joined.
        \item As soon as the iPad launched, ``the coordinated pressure on Amazon 
        began at once.''\footnote{p. 83.}
        \item Jobs acknowledged to a reporter that the goal was to eliminate 
        competition from Amazon and raise ebook prices.\footnote{p. 86.}
        \item Soon, the ebooks saw ``sudden and uniform price increases'' across 
        the market.\footnote{p. 94.} ``The publisher defendants used the change 
        to an agency method for distributing their e-books as an opportunity to 
        raise the prices for their e-books across the board.''\footnote{p. 96.}
        \item The publishers lost sales as a result of the price increases, and 
        consumers suffered because they had to pay more for ebooks, buy cheaper 
        ebooks, or defer purchases entirely.\footnote{p. 98.}
    \end{enumerate}
    \item Holding:
    \begin{enumerate}
        \item The Sherman Act only outlaws \emph{unreasonable} restraints on 
        trade. Some offenses are per se anticompetitive, while others are 
        analyzed under the rule of reason. Courts have to weigh the competitive 
        effects of the agreement.\footnote{pp. 105--106.}
        \item Horizontal price fixing is illegal per se.\footnote{p. 107.}
        \item A price-fixing agreement requires a meeting of the 
        minds.\footnote{p. 109.}
        \item The court here found ``overwhelming evidence'' of a price-fixing 
        conspiracy.\footnote{p. 113.} Plaintiffs ``have proven a per se 
        violation of the Sherman Act.''\footnote{p. 120.}
        \item Vertical players can be liable per se for facilitating a 
        horizontal agreement. \emph{Toys R Us}.
    \end{enumerate}
    \item Apple's arguments: p. 122 ff.
\end{enumerate}
