\section{Smartphone Wars}

\subsection{Standard-Essential Patents}

\begin{enumerate}
    \item Standards-setting organizations exist to eliminate competing 
    solutions to technological problems.
    \item Solutions are selected on (1) the merits and (2) their availability 
    and cost, which depends on patent encumbrances. So, they demand that 
    patents be declared, and they demand a declaration of the conditions under 
    which the patent holder would be willing to license it.
    \item FRAND: fair, reasonable, and nondiscriminatory licensing terms.
    \item FRAND commitments do not provide clear guidance on the parameters of 
    a reasonable and nondiscriminatory license, so \textbf{Hold-up} is a risk, 
    particularly since SSOs adopt the SEPs before royalty rates are 
    negotiated. The FTC has pushed for various remedies, including use of the 
    \emph{eBay} factors for injunctions instead of automatic injunctions, and 
    a broad interpretation of ITC \S\ 337.
\end{enumerate}

\subsection{Broadcom Corp. v. Qualcomm Inc.}

Exclusionary conduct: making and breaking a FRAND commitment.

\begin{enumerate}
    \item Broadcom alleged that Qualcomm, by acting deceptively towards private 
    standards organizations and predatorily acquiring a rival, violated \S\S\ 
    1 and 2 of the Sherman Act and \S\ 7 of the Clayton Act.\footnote{303.}
    \item The district court granted Qualcomm's 12(b)(6) motion.
    \item Held: Broadcom stated claims for monopolization and attempted 
    monopolization under \S\ 2, but that it lacked standing because it does 
    not compete or intend to compete in the market.
    \item The technology at issue was Qualcomm's WCDMA, part of the 3G system 
    known as the UMTS standard.\footnote{303-04.}
    \item Claim 1: Monopolization under \S\ 2.
    \begin{enumerate}
        \item The district court erred in dismissing. Abuse of a private 
        standard-setting process \emph{does} state a claim under antitrust 
        law.\footnote{306--17.}
    \end{enumerate}
    \item Claim 2: Attempted monopolization.
    \begin{enumerate}
        \item The district court erred in dismissing.\footnote{317--18.}
    \end{enumerate}
    \item Claim 7: Monopoly maintenance.
    \begin{enumerate}
        \item Dismissal was proper.\footnote{319--21.}
    \end{enumerate}
\end{enumerate}
